Wetting of the tarsal adhesive fluid controls underwater adhesion in ladybug beetles

### 2021-05-11 source/tex/ladybug_adhesion.tex
final formatting edits for submission to JEB

\begin{document}
\begin{abstract}

@@

While most of the studies on insect adhesion focused on terrestrial species, underwater insect attachment is much rarer and has been relatively unexplored. Some aquatic insects like diving beetles \cite{Chen:2014} or midge larva \cite{Kang:2020} use suction cups to adhere to surfaces. However, underwater adhesion mediated by secreted liquids require the displacement of the water at the interface first and a spreading of the fluid on the substrate. One relatively simple approach is to use an air bubble around the adhesive organs similar to the air bubbles many secondary aquatic insects and spiders carry on their body for breathing underwater \cite{Seymour:2013}. This has been shown in a recent study by \citet{RN87} that that female terrestrial green dock beetles \emph{Gastrophysa viridula} can attach quite well to surfaces underwater by using such an air bubble. Their naturally hydrophobic tarsal hairs trap the bubble around the pads when being submerged underwater, which de-wets the surface on contact. It has been hypothesised that a combination of capillary forces due the air bubble and hair secretions within the de-wetted area results in its adhesion underwater. However, it remained unclear if an air bubble is necessary for adhesion and what, if any, contribution it has to the adhesive force. The oily tarsal adhesive fluid found in insects alone might be sufficient in creating the necessary capillary adhesion even without a bubble, given that the fluid remains on the hair tips when submerged. In ladybug beetles, the tip of each seta secretes approximately one femtoliter of tarsal adhesive fluid by each step \cite{RN108}. The fluid's chemical composition in green dock beetles was identified to be an oil-containing mixture of mostly long chain hydrocarbons\cite{RN96} with traces
of triglycerides, fatty acids and cholesterol in ladybirds \emph{Hemisphaerota cyanea} \cite{RN221} and \emph{Epilachna vingtipunctuata} \cite{RN222}, rendering it immiscible with water.

@@

 %The beetle's leg was constrained similar to the method described by Bullock et al \cite{RN19}.
An individual beetle was first carefully anaesthetised using small amounts of CO\textsubscript{2} sublimating
from a piece of dry ice and then glued with a small dollop of epoxy glue on its elytra to the underside of a heavy steel ball. The ball was held in a bracket which allowed free rotational movement of the ball in each direction, thus helping to align the suspended beetle over the substrate (see Fig.~\ref{fig:Setup}). The bracket with the ball and the beetle could be further positioned by manual micro-manipulators in all three axes before the experiments. One front left leg was carefully fixed at its tibia to a piece of soft solder wire coming off the steel ball using Blu Tack (Bostik Ltd., U.K.), allowing us to further align the leg to the substrate. Each leg of a male ladybug beetle has two hairy adhesive pads. For the test, we only allowed the distal pad to come into contact with the substrate thus minimising partial or bad contact of the other one. The distal pad was thus restrained by fixing its dorsal side to the wire using Blu Tack. The claws on the leg were also fixed to the wire using epoxy glue to prevent any further movement and to prevent the claws from touching the substrate (Fig.~\ref{fig:Setup} top-left inset). Care was taken to ensure the glue doesn't contaminate the rest of the tarsomeres. A small piece of non-sticky Teflon tape helped to keep other legs tucked close to the body and avoided any interference during the adhesion test. After the measurements, the beetle was freed by carefully removing the epoxy glue and Blu Tack without harming it and set free.

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.5]{Figure1-Setup_schematic}
\par\end{centering}
\caption{\label{fig:Setup} Adhesion test setup (see text for details). Top-left inset shows a magnified cartoon of the beetle's leg constrained to a solder wire (grey) using Blu Tack (blue) and epoxy glue (orange). The recorded force data and
contact area of a distal pad are shown in the bottom-right plot, in which, the shaded
regions from left to right represent the distinct movement sequence: approach, lateral pull, approach, pause and retract, respectively. Negative force values
represent attraction and the minimum force peak during the final retraction
step is the adhesion force used for further analysis. An animated version of a typical force recording is available in the Supplementary Material (S2-S4).}
\end{figure}

@@

Data sets were compared for statistical differences using two-way ANOVA analysis, with \emph{contact mode} and \emph{substrate chemistry} as the categorical variables and \emph{adhesion force} as the dependant variable. Pairwise
Student t-test were done for post-hoc analysis and their corresponding p-value and Common Language
Effect Size (CLES) are reported. Shapiro-Wilk test was done for each
data set to verify a normal distribution of its residuals and Levene's
test was done to check for variance homogeneity, to validate the t-test
assumptions. Bonferroni's correction was used to account for multiple
comparison between groups.

@@

The surface chemistry was characterised by dynamic contact angle
measurements, performed with a contact angle goniometer (\emph{OCA 35, DataPhysics Instruments GmbH, Germany})
The substrate's wetting towards a polar (Milli-Q water) and a non-polar (n-hexadecane) liquid was tested. Advancing
and receding contact angles were measured for a maximum drop volume
of 10 \textmu l and with 0.5 \textmu l s\protect\textsuperscript{-1} flow rate (Table \ref{tab:Contact-Angles}).

\begin{table}[H]
\centering{}%
\begin{tabular}{|c|c|c|c|}
\hline 
Substrate & Liquid & \ensuremath{\theta}\textsubscript{A} & \ensuremath{\theta}\textsubscript{R}\tabularnewline
\hline 
\hline 
\multirow{2}{*}{Glass} & Water & 63\ensuremath{\pm}5\textdegree{} & 20\ensuremath{\pm}2\textdegree{}\tabularnewline
\cline{2-4} \cline{3-4} \cline{4-4} 
 & n-Hexadecane & <10\textdegree{} & <10\textdegree{}\tabularnewline
\hline 
\multirow{2}{*}{PFOTS} & Water & 122\ensuremath{\pm}1\textdegree{} & 93\ensuremath{\pm}2\textdegree{}\tabularnewline
\cline{2-4} \cline{3-4} \cline{4-4} 
 & n-Hexadecane & 88\ensuremath{\pm}2\textdegree{} & 56\ensuremath{\pm}5\textdegree{}\tabularnewline
\hline 
\end{tabular}\caption{Dynamic contact angles (Mean \ensuremath{\pm} SD, n = 3) of Milli-Q water and n-hexadecane on the different test substrates. \label{tab:Contact-Angles}}
\end{table}

@@

In air, adhesion forces of the distal pad of ladybug beetle against
glass and PFOTS were similar, i.e. no significant differences were
detected (Figure \ref{fig:Effect-of-contact} and Table S1.1).
In contrast, the underwater adhesion on a PFOTS surface was significantly
larger than on glass (p < 0.001). This stronger adhesion on PFOTS
was observed both in the presence and absence of a trapped bubble.
In both cases, the adhesion force reached similar values as in air. In contrast,
on glass, adhesion underwater was significantly reduced when compared
to dry conditions, irrespective of the presence of a trapped bubble
(p $\leq$ 0.002). In the presence of a bubble, underwater adhesion on glass was slightly higher (CLES = 0.84, p = 0.07).

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{Figure2-Expt_effect_of_contact}\caption{\label{fig:Effect-of-contact}Box-and-whisker plot showing
 adhesion force measurements of ladybug beetle's (\emph{Coccinella
septempuctata}) distal pad on untreated hydrophilic glass (blue) and hydrophobic PFOTS coated glass (red) substrates in air and underwater
conditions (n=5 per box). The two modes of contact during underwater experiments
are represented separately: \textquotedblleft\emph{bubble}\textquotedblright{}
and \textquotedblleft\emph{no bubble}\textquotedblright . The small diamonds represent outlier points. Crosses
represent theoretical predictions (see text) of adhesion force, while, circles
represent the contribution of the bubble itself, calculated from the
capillary bridge model. In the model, hair diameter = 4 \textmu m,
pad diameter = 200 \textmu m, hair length = 40 \textmu m,
N\protect\textsubscript{hairs} = 500, V\protect\textsubscript{fluid}
= 4.2 fL and V\protect\textsubscript{bubble} = 1.2 nL. Interfacial
tension of the tarsal adhesive fluid in air and water were assumed to be
24 mN m\protect\textsuperscript{-1} and 48 mN m\protect\textsuperscript{-1} respectively and water surface tension is 72 mN m\protect\textsuperscript{-1}.}
\end{figure}

Apart from the three depicted contact modes, we observed an additional
fourth mode which occurred in roughly 25\% of our underwater experiments using degassed water.
In this scenario, the ventral view recordings show that none of the hairs
appear to contact well with either glass or PFOTS substrate (supplementary video S3), unlike the other
three contact modes (supplementary video S2). This ``bad contact'' scenario only happened underwater and shows no adhesion 
with either glass or PFOTS substrate. While it was not completely
clear why such a contact occurs, there can be two possible reasons.
First, the hairs could get bundled due to a small air meniscus within
the hairs, resulting in their disorientation. Second, a thin water
layer at the substrate interface might not be drained out to allow
the hairs to make contact with the substrate, resulting in a loss
of adhesion.

@@

The capillary force, $f$, is the sum of two contributions: surface
tension and Laplace pressure (equation S1.1). Force
versus distance for a single capillary bridge was calculated by Surface
Evolver simulations\cite{RN206}  (described in supplementary section S1.1),
and used to obtain $F_{net}$ as a function of $d$ for each mode
of contact. The adhesion force of the complete hairy pad system was
then obtained from the minima of $F_{net}$, where negative force
values represent attraction.

@@

\subcaption{Force contributions}%
\end{minipage}\caption{\label{fig:Single-bridge}Simulation of normalised capillary force of a single liquid
bridge in contact with a substrate and pinned to a circular perimeter
on top. Fluid size parameter, $\phi_{f}=2$. Negative force values
represents attraction. a) Force-distance curves are shown for different
contact angles of the liquid with the substrate. b) Adhesion forces,
calculated from the minima of the corresponding force-distance curves,
are plotted as a function of contact angle with the substrate, together
with its Laplace and surface tension components (equation S1.1).
Simulation snapshots of the liquid meniscus corresponding to angles
6\textdegree{} and 150\textdegree{} are depicted.}
\end{figure}

@@

Our experiments demonstrate that the ladybug beetle can attach underwater
to a hydrophobic substrate even without a bubble trapped around its
tarsal hairs. A previous study\cite{RN87} proposed that an air bubble
is necessary for underwater attachment in terrestrial beetles. This is, however,
only true for hydrophilic substrates, where a trapped air bubble can facilitate
underwater adhesion due to the hairs making contact in a de-wetted
environment. For a hydrophobic substrate, the adhesion is similar
regardless of whether the contact occurs in air or underwater conditions,
with or without a trapped bubble. Our theoretical calculations further
show that the bubble by itself has a negligible capillary contribution
to the net underwater adhesion of the pad. Direct force measurement
of a single similarly sized bubble making contact with a hydrophobic
substrate shows a maximum adhesion less than 50 \textmu N, which
further validates that the bubble's contribution is insignificant
(Figure S1.3).

@@

We have so far limited our analysis to only smooth substrates. Of course insects
have to cope with all kinds of surfaces including rough ones.
Previous studies \cite{RN136} have shown that substrate roughness
is a more dominant parameter than substrate chemistry in controlling
ladybug beetle traction force. Future work will explore how roughness can impact
the net capillary force also in wet and submerged conditions. We have limited our analysis only to normal adhesive forces, but insects like beetles in general rely on friction or shear forces during locomotion. Friction force usually correlates directly with the normal force, which is probably why previously reported shear adhesion forces of the dock beetle \cite{RN87} follow a similar qualitative trend as our normal adhesion force measurements on the ladybug beetle in both air and underwater conditions. However, the details of the interplay between friction and normal adhesion forces in animals is an open question and is beyond the scope of this paper.

@@

\section{Acknowledgement}

We are grateful to Eduard Arzt and Ren\textcyr{\`\cyre} Hensel (Leibniz Institute for New Materials, Saarbr\"{u}cken, Germany)  for
fruitful discussions. This work was supported by the \emph{Deutsche Forschungsgemeinschaft}
(Grant number: PI 1351/2-1) and the \emph{Max Planck Graduate Center} with the \emph{Johannes Gutenberg-Universit\"{a}t Mainz} (MPGC).

>>>>

\begin{document}

\section*{Summary statement}

Fluid-mediated adhesion seen in animals such as ladybug beetles allow them to attach to surfaces not just in air but also in underwater conditions.

\begin{abstract}

@@

While most of the studies on insect adhesion focused on terrestrial species, underwater insect attachment is much rarer and has been relatively unexplored. Some aquatic insects like diving beetles \cite{Chen:2014} or midge larva \cite{Kang:2020} use suction cups to adhere to surfaces. However, underwater adhesion mediated by secreted liquids require the displacement of the water at the interface first and a spreading of the fluid on the substrate. One relatively simple approach is to use an air bubble around the adhesive organs similar to the air bubbles many secondary aquatic insects and spiders carry on their body for breathing underwater \cite{Seymour:2013}. This has been shown in a recent study by \citet{RN87} that female terrestrial green dock beetles \emph{Gastrophysa viridula} can attach quite well to surfaces underwater by using such an air bubble. Their naturally hydrophobic tarsal hairs trap the bubble around the pads when being submerged underwater, which de-wets the surface on contact. It has been hypothesised that a combination of capillary forces due the air bubble and hair secretions within the de-wetted area results in its adhesion underwater. However, it remained unclear if an air bubble is necessary for adhesion and what, if any, contribution it has to the adhesive force. The oily tarsal adhesive fluid found in insects alone might be sufficient in creating the necessary capillary adhesion even without a bubble, given that the fluid remains on the hair tips when submerged. In ladybug beetles, the tip of each seta secretes approximately one femtoliter of tarsal adhesive fluid by each step \cite{RN108}. The fluid's chemical composition in green dock beetles was identified to be an oil-containing mixture of mostly long chain hydrocarbons\cite{RN96} with traces
of triglycerides, fatty acids and cholesterol in ladybirds \emph{Hemisphaerota cyanea} \cite{RN221} and \emph{Epilachna vingtipunctuata} \cite{RN222}, rendering it immiscible with water.

@@

 %The beetle's leg was constrained similar to the method described by Bullock et al \cite{RN19}.
An individual beetle was first carefully anaesthetised using small amounts of CO\textsubscript{2} sublimating
from a piece of dry ice and then glued with a small dollop of epoxy glue on its elytra to the underside of a heavy steel ball. The ball was held in a bracket which allowed free rotational movement of the ball in each direction, thus helping to align the suspended beetle over the substrate (see Fig.~\ref{fig:Setup}). The bracket with the ball and the beetle could be further positioned by manual micro-manipulators in all three axes before the experiments. One front left leg was carefully fixed at its tibia to a piece of soft solder wire coming off the steel ball using Blu Tack (Bostik Ltd., U.K.), allowing us to further align the leg to the substrate. Each leg of a male ladybug beetle has two hairy adhesive pads. For the test, we only allowed the distal pad to come into contact with the substrate thus minimising partial or bad contact of the other one. The distal pad was thus restrained by fixing its dorsal side to the wire using Blu Tack. The claws on the leg were also fixed to the wire using epoxy glue to prevent any further movement and to prevent the claws from touching the substrate (Fig.~\ref{fig:Setup} top-left inset). Care was taken to ensure the glue doesn't contaminate the rest of the tarsomeres. A small piece of non-sticky Teflon tape helped to keep the other legs tucked close to the body and avoided any interference during the adhesion test. After the measurements, the beetle was freed by carefully removing the epoxy glue and Blu Tack without harming it and set free.

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.5]{Figure1-Setup_schematic}
\par\end{centering}
\caption{\label{fig:Setup} Adhesion test setup (see text for details). Top-left inset shows a magnified cartoon of the beetle's leg constrained to a solder wire (grey) using Blu Tack (blue) and epoxy glue (orange). The recorded force data and
contact area of a distal pad are shown in the bottom-right plot, in which, the shaded
regions from left to right represent the distinct movement sequence: approach, lateral pull, approach, pause and retract, respectively. Negative force values
represent attraction and the minimum force peak during the final retraction
step is the adhesion force used for further analysis. An animated version of a typical force recording is available in the supplementary material (see Movie1).}
\end{figure}

@@

Data sets were compared for statistical differences using two-way ANOVA analysis, with \emph{contact mode} and \emph{substrate chemistry} as the categorical variables and \emph{adhesion force} as the dependant variable. Pairwise
Student t-test were done for post-hoc analysis and their corresponding p-value and Common Language
Effect Size (CLES) are reported. Shapiro-Wilk test was done for each
data set to verify a normal distribution of its residuals and Levene's
test was done to check for variance homogeneity, to validate the ANOVA 
assumptions. Bonferroni's correction was used to account for multiple
comparison between groups.

@@

The surface chemistry was characterised by dynamic contact angle
measurements, performed with a contact angle goniometer (\emph{OCA 35, DataPhysics Instruments GmbH, Germany})
The substrate's wetting towards a polar (Milli-Q water) and a non-polar (n-hexadecane) liquid was tested. Advancing
and receding contact angles were measured for a maximum drop volume
of 10 \textmu l and with 0.5 \textmu l s\protect\textsuperscript{-1} flow rate (see supplementary section S2).

@@

In air, adhesion forces of the distal pad of ladybug beetle against
glass and PFOTS were similar, i.e. no significant differences were
detected (Figure \ref{fig:Effect-of-contact} and supplementary section S3).
In contrast, the underwater adhesion on a PFOTS surface was significantly
larger than on glass (p < 0.001). This stronger adhesion on PFOTS
was observed both in the presence and absence of a trapped bubble.
In both cases, the adhesion force reached similar values as in air. In contrast,
on glass, adhesion underwater was significantly reduced when compared
to dry conditions, irrespective of the presence of a trapped bubble
(p $\leq$ 0.002). In the presence of a bubble, underwater adhesion on glass was slightly higher (CLES = 0.84, p = 0.07).

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{Figure2-Expt_effect_of_contact}\caption{\label{fig:Effect-of-contact}Box-and-whisker plot showing
 adhesion force measurements of ladybug beetle's (\emph{Coccinella
septempuctata}) distal pad on untreated hydrophilic glass (blue) and hydrophobic PFOTS coated glass (red) substrates in air and underwater
conditions (n=5 per box). The two modes of contact during underwater experiments
are represented separately: \textquotedblleft\emph{bubble}\textquotedblright{}
and \textquotedblleft\emph{no bubble}\textquotedblright . The small diamonds represent outlier points. Crosses
represent theoretical predictions of adhesion force, while, circles
represent the contribution of the bubble itself, calculated from the
capillary bridge model (see text). In the model, hair diameter = 4 \textmu m,
pad diameter = 200 \textmu m, hair length = 40 \textmu m,
N\protect\textsubscript{hairs} = 500, V\protect\textsubscript{fluid}
= 4.2 fL and V\protect\textsubscript{bubble} = 1.2 nL. Interfacial
tension of the tarsal adhesive fluid in air and water were assumed to be
24 mN m\protect\textsuperscript{-1} and 48 mN m\protect\textsuperscript{-1} respectively and water surface tension is 72 mN m\protect\textsuperscript{-1}.}
\end{figure}

Apart from the three depicted contact modes, we observed an additional
fourth mode which occurred in roughly 25\% of our underwater experiments using degassed water.
In this scenario, the ventral view recordings show that none of the hairs
appear to contact well with either glass or PFOTS substrate (supplementary data: Movie2), unlike the other
three contact modes (supplementary data: Movie1). This ``bad contact'' scenario only happened underwater and shows no adhesion 
with either glass or PFOTS substrate. While it was not completely
clear why such a contact occurs, there can be two possible reasons.
First, the hairs could get bundled due to a small air meniscus within
the hairs, resulting in their disorientation. Second, a thin water
layer at the substrate interface might not be drained out to allow
the hairs to make contact with the substrate, resulting in a loss
of adhesion.

@@

The capillary force, $f$, is the sum of two contributions: surface
tension and Laplace pressure. Force
versus distance for a single capillary bridge was calculated by Surface
Evolver simulations\cite{RN206}  (see supplementary section S1 for details),
and used to obtain $F_{net}$ as a function of $d$ for each mode
of contact. The adhesion force of the complete hairy pad system was
then obtained from the minima of $F_{net}$, where negative force
values represent attraction.

@@

\subcaption{Force contributions}%
\end{minipage}\caption{\label{fig:Single-bridge}Simulation of normalised capillary force of a single liquid
bridge in contact with a substrate and pinned to a circular perimeter
on top. Fluid size parameter, $\phi_{f}=2$. Negative force values
represents attraction. a) Force-distance curves are shown for different
contact angles of the liquid with the substrate. b) Adhesion forces,
calculated from the minima of the corresponding force-distance curves,
are plotted as a function of contact angle with the substrate, together
with its Laplace and surface tension components (equation S1 in supplementary section).
Simulation snapshots of the liquid meniscus corresponding to angles
6\textdegree{} and 150\textdegree{} are depicted.}
\end{figure}

@@

Our experiments demonstrate that the ladybug beetle can attach underwater
to a hydrophobic substrate even without a bubble trapped around its
tarsal hairs. A previous study\cite{RN87} proposed that an air bubble
is necessary for underwater attachment in terrestrial beetles. This is, however,
only true for hydrophilic substrates, where a trapped air bubble can facilitate
underwater adhesion due to the hairs making contact in a de-wetted
environment. For a hydrophobic substrate, the adhesion is similar
regardless of whether the contact occurs in air or underwater conditions,
with or without a trapped bubble. Our theoretical calculations further
show that the bubble by itself has a negligible capillary contribution
to the net underwater adhesion of the pad. Direct force measurement
of a single similarly sized bubble making contact with a hydrophobic
substrate shows a maximum adhesion less than 50 \textmu N, which
further validates that the bubble's contribution is insignificant
(see supplementary section S4).

@@

We have so far limited our analysis to only smooth substrates. Of course insects
have to cope with all kinds of surfaces including rough ones.
Previous studies \cite{RN136} have shown that substrate roughness
is a more dominant parameter than substrate chemistry in controlling
ladybug beetle traction force. Future work will explore how roughness can impact
the net capillary force also in wet and submerged conditions. In our study, we have only considered normal adhesive forces, but insects like beetles in general rely on friction or shear forces during locomotion. Friction force usually correlates directly with the normal force, which is probably why previously reported shear adhesion forces of the dock beetle \cite{RN87} follow a similar qualitative trend as our normal adhesion force measurements on the ladybug beetle in both air and underwater conditions. However, the details of the interplay between friction and normal adhesion forces in animals is an open question and is beyond the scope of this paper.

@@

\section{Acknowledgements}

We are grateful to Prof. Dr. Eduard Arzt and Dr. Ren\textcyr{\`\cyre} Hensel (Leibniz Institute for New Materials, Saarbr\"{u}cken, Germany)  for fruitful discussions.

\section{Competing interests}
The authors declare no competing or financial interests.

%\section{Data availability}
%Data are available from the Dryad digital repository. %CHECK THIS%

\section{Funding}
This work was supported by the \emph{Deutsche Forschungsgemeinschaft}
(Grant number: PI 1351/2-1) and the \emph{Max Planck Graduate Center} with the \emph{Johannes Gutenberg-Universit\"{a}t Mainz} (MPGC).
